---
title: "homework 2"
author: "lauren franks, lef2147"
date: "10/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Libraries
```{r, message=FALSE}
library(readxl)
library(tidyr)
library(dplyr)
library(readr)
library(tidyverse)
```


Question 1:
Reading and Cleaning Dataset
```{r}
trash_data = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", cell_cols("A:N"), ) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls) %>% 
      as.integer()
  )

precip_data_2017 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(
    year = 2017
  ) %>% 
  drop_na(total)

precip_data_2018 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(
    year = 2018,
  ) %>% 
  drop_na(total)

```

Combine precipitation datasets & change number to month
```{r}
precip_data = bind_rows(precip_data_2017, precip_data_2018) %>% 
  mutate(
    month = month.name[month]
  )
```
The dataset `trash_data` has `r nrow(trash_data)` observations and `r ncol(trash_data)` variables. There were on average `r mean(pull(trash_data, sports_balls))` sports balls removed per day, and an average of `r mean(pull(trash_data, weight_tons))` tons of trash removed each day. There was total of `r sum(pull(precip_data_2017, total))` inches of rain in 2017 and `r sum(pull(precip_data_2018, total))` inches in 2018.


Question 2:
Import and Clean Data
president: 1 = republican, 0 = democrat
```{r, message=FALSE}
pols_month = read_csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(col = mon, into = c("year", "month", "day"), "-") %>% 
  mutate(
    president = prez_gop,
    month = str_remove(month, "^0+"),
    month = as.integer(month),
    month = month.abb[month],
    year = as.integer(year)
  ) %>% 
  subset(select = -c(prez_gop, prez_dem, day))
```

Tidying snp dataset
```{r, message=FALSE}
snp_data = read_csv("./fivethirtyeight_datasets/snp.csv") %>% 
  separate(col = date, into = c("month", "day", "year"), "/", convert = TRUE) %>% 
  subset(select = -day) %>% 
  arrange(year, month) %>% 
  select(year, month, close) %>% 
  mutate(
    month = month.abb[month]
  )
```

Tidying unemployment data 
```{r, message=FALSE}
unemploy_data = read_csv("./fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemp_percent"
  ) %>% 
  janitor::clean_names()
```

Merging datasets
```{r, message=FALSE}
fivethirtyeight = left_join(pols_month, snp_data, key = c("year", "month"))
fivethirtyeight = left_join(fivethirtyeight, unemploy_data, key = c("year", "month"))
```
The three datasets were joined together to create one cohesive dataset. The first dataframe, `pols_month`, contained information regarding the national politicians categorized by democrats and republicans. There are `r nrow(pols_month)` observations and `r ncol(pols_month)` variables. The second dataframe, `snp_data`, contained information regarding the closing value of the S&P stock market to give an overall measure of stock on that date. There are `r nrow(snp_data)` observations. The third dataframe, `unemploy_data`, gives the unemployment percentage for the assocaited date. There are `r nrow(unemploy_data)` observations. The resulting dataframe of the three dataframes combined, `fivethirtyeight`, has `r nrow(fivethirtyeight)` observations and `r ncol(fivethirtyeight)` variables. The years range from `r range(pull(fivethirtyeight, year))`, however certain data is not available for all years, e.g. 1947 does not have unemployment percentage (unemp_percent) and SNP close value (close). 


Question 3
Loading data
```{r, message=FALSE}
baby_names = read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% #unique() %>% 
  mutate(
    gender = as.factor(gender),
    ethnicity = as.factor(ethnicity),
  ) %>% unique()

baby_names = data.frame(lapply(baby_names, function(childs_first_name) {
  if (is.character(childs_first_name)) return(toupper(childs_first_name))
  else return(childs_first_name)
}))
```

Produce a table showing the rank in pop of the name "Olivia" as a female name over time, include ethinicty and year.
```{r}
baby_names %>% filter(childs_first_name == "OLIVIA") %>% 
  subset(select = -c(gender, childs_first_name, count)) %>% 
  mutate(
    rank = as.integer(rank)
  ) %>% 
  pivot_wider(
    names_from = ethnicity,
    values_from = rank,
    values_fill = list(rank = 0)
  ) %>% knitr::kable()
```

Finding most popular male name over time
```{r}
male_pop = baby_names %>% filter(gender == "MALE") %>% 
  group_by(childs_first_name) %>%
  summarise(count1 = sum(count)) %>% 
  mutate(
    count1 = as.integer(count1)
  )

baby_names %>% filter(childs_first_name == "ETHAN") %>% 
  subset(select = -c(gender, childs_first_name, count)) %>% 
  mutate(
    rank = as.integer(rank)
  ) %>% 
  pivot_wider(
    names_from = ethnicity,
    values_from = rank,
    values_fill = list(rank = 0)
  ) %>% knitr::kable()
```


Male, white non-hispanic children born in 2016, produce a scatterplot showing the number of children with a name (y axis) against the rank in popularity of the name (x axis)
```{r}
subset_data = baby_names %>% 
  filter(gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == "2016")

subset_data %>% 
  ggplot(aes(rank, count)) + 
  geom_point() +
  labs(
    title = "Number of Children with a name against the name's rank in popularity"
  )
```



